---
title: "Ask a Manager Survey"
description: |
  2021-05-18
author:
  - name: Vincent Manzanilla
    url: https://github.com/vincentmanz
    affiliation: BaseClear B.V / Datalabs
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 200
  chunk_output_type: console
---



The data this week comes from the Ask a Manager Survey. H/t to Kaija Gahm for sharing it as an issue!

    The salary survey a few weeks ago got a huge response — 24,000+ people shared their salaries and other info, which is a lot of raw data to sift through. Reader Elisabeth Engl kindly took the raw data and analyzed some of the trends in it and here’s what she found. (She asked me to note that she did this as a fun project to share some insights from the survey, rather than as a paid engagement.)

    This data does not reflect the general population; it reflects Ask a Manager readers who self-selected to respond, which is a very different group (as you can see just from the demographic breakdown below, which is very white and very female).
 
Elisabeth Engl prepped some plots for the [Ask a Manager blog](https://www.askamanager.org/2021/05/some-findings-from-24000-peoples-salaries.html) using this data.

```{r knitr, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, fig.showtext = T, fig.retina = 1)
```

```{r setup}
library(tidyverse)
library(ggplot2)
library(hrbrthemes)
library(kableExtra)
library(showtext)
library(ggtext)
library(patchwork)

font_add_google("Oswald", "oswald")
font_add_google("Roboto Condensed", "roboto condensed")
font_add_google("Share Tech Mono", "techmono")

showtext_opts(dpi = 320)
showtext_auto(enable = TRUE)


## ggplot theme updates
source(here::here("theme", "tidy_grey.R"))
```


# Get the data

```{r get the data}
# Get the Data

# Read in with tidytuesdayR package 
# Install from CRAN via: install.packages("tidytuesdayR")
# This loads the readme and all the datasets for the week of interest

# Either ISO-8601 date or year/week works!

tuesdata <- tidytuesdayR::tt_load('2021-05-18')
tuesdata <- tidytuesdayR::tt_load(2021, week = 21)

survey <- tuesdata$survey

# Or read in the data manually
#survey <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-18/survey.csv')

```


```{r}
#from https://github.com/richardvogg/tidy-tuesdays/blob/master/2021/Week%2021%20-%20salary%20survey/salary_survey.R
survey_filtered <- survey %>%
  mutate(Continent = str_remove_all(country, "[[:punct:]]") %>%
           tolower() %>%
           str_replace("^australi.*","Oceania") %>%
           str_replace("^argentina*","South America") %>%
           str_replace("^austria.*","Europe") %>%
           str_replace("^bangladesh","Asia") %>%
           str_replace("^belgium","Europe") %>%
           str_replace("^bermuda","North America") %>%
           str_replace("^brasil|^brazil","South America") %>%
           str_replace("^britain","Europe") %>%
           str_replace("^bulgaria","Europe") %>%
           str_replace("^california","North America") %>%
           str_replace("^cambodia","Asia") %>%
           str_replace("^canad.*|canda.*|csnada.*","North America") %>%
           str_replace("^catalonia","Europe") %>%
           str_replace("^cayman islands","North America") %>%
           str_replace("^chile","South America") %>%
           str_replace("^china","Asia") %>%
           str_replace("^colombia","South America") %>%
           str_replace("^afghanistan","Asia") %>%
           str_replace("^africa","Africa") %>%
           str_replace("^america","South America") %>%
           str_replace("^congo","Africa") %>%
           str_replace("^costa rica","South America") %>%
           str_replace("^cote divoire","Africa") %>%
           str_replace("^croatia","Europe") %>%
           str_replace("^cuba","North America") %>%
           str_replace("^czech.*","Europe") %>%
           str_replace("^danmark|^denmark","Europe") %>%
           str_replace("^ecuador","South America") %>%
           str_replace("^england.*|^englang","Europe") %>%
           str_replace("^eritrea","Africa") %>%
           str_replace("^estonia","Europe") %>%
           str_replace("^eu.*","Europe") %>%
           str_replace("europe|EUraine","Europe") %>%
           str_replace("^finland","Europe") %>%
           str_replace("^france","Europe") %>%
           str_replace("^germany","Europe") %>%
           str_replace("^ghana","Africa") %>%
           str_replace("^great britain","Europe") %>%
           str_replace("^greece","Europe") %>%
           str_replace("^hartford","Europe") %>%
           str_replace("^hong kong","Asia") %>%
           str_replace("^hungary","Europe") %>%
           str_replace("^india","Asia") %>%
           str_replace("^indonesia","Asia") %>%
           str_replace("^ireland","Europe") %>%
           str_replace("^bangladesh","Asia") %>%
           str_replace("^bangladesh","Asia") %>%
           str_replace("isle of man","Europe") %>%
           str_replace("^israel","Asia") %>%
           str_replace("^is.*","North America") %>%
           str_replace("^isa.*","North America") %>%
           str_replace("^italy.*","Europe") %>%
           str_replace("^jamaica","North America") %>%
           str_replace("^japan.*","Asia") %>%
           str_replace("^jersey.*","Europe") %>%
           str_replace("^jordan","Asia") %>%
           str_replace("^kenya","Africa") %>%
           str_replace("^kuwait","Asia") %>%
           str_replace("^latvia","Europe") %>%
           str_replace("^lithuania","Europe") %>%
           str_replace("^luxemb.*","Europe") %>%
           str_replace("^mainland china","Asia") %>%
           str_replace("^malaysia","Asia") %>%
           str_replace("^malta","Europe") %>%
           str_replace("^mexico|^méxico","North America") %>%
           str_replace("^morocco","Africa") %>%
           str_replace("^nederland.*|^nl|^netherlands|^the netherlands","Europe") %>%
           str_replace("^new zealand.*|^nz","Oceania") %>%
           str_replace("^nigeria","Europe") %>%
           str_replace("northern ireland.*","Europe") %>%
           str_replace("norway","Europe") %>%
           str_replace("^pakistan","Asia") %>%
           str_replace("^panamá","North America") %>%
           str_replace("^philippines|^remote philippines","Asia") %>%
           str_replace("^poland","Europe") %>%
           str_replace("^portugal","Europe") %>%
           str_replace("^puerto rico","North America") %>%
           str_replace("^qatar","Asia") %>%
           str_replace("^romania","Europe") %>%
           str_replace("^russia","Asia") %>%
           str_replace("^rwanda","Africa") %>%
           str_replace("^san francisco","North America") %>%
           str_replace("^saudi arabia","Asia") %>%
           str_replace("^scotland.*","Europe") %>%
           str_replace("^serbia","Europe") %>%
           str_replace("^sierra leone","Africa") %>%
           str_replace("^singapore","Asia") %>%
           str_replace("^slovakia","Europe") %>%
           str_replace("^slovenia","Europe") %>%
           str_replace("^somalia","Africa") %>%
           str_replace("^south africa","Africa") %>%
           str_replace("^south korea","Asia") %>%
           str_replace("^spain","Europe") %>%
           str_replace("^sri lanka","Asia") %>%
           str_replace("^sweden","Europe") %>%
           str_replace("^switzerland","Europe") %>%
           str_replace("^taiwan","Asia") %>%
           str_replace("^thailand","Asia") %>%
           str_replace("^the bahamas","North America") %>%
           str_replace("^the united states.*|^the us.*|^u s.*|^ua.*|^uniited states.*|^unite states.*|^united  states.*|^united sates.*|^united state.*|^united stares.*|^united stat.*|^united sttes.*|^unitedstates.*|^uniteed states.*|^unitef stated.*|^uniter statez.*|^unites states|^unitied states.*|^uniyed states.*|^uniyes states.*|^unted states.*|^untied states.*|^united statss.*|^united stattes.*|^united statues.*|^united status.*|^united statws.*|^unitied states.*","North America") %>%
           str_replace("^turkey","Asia") %>%
           str_replace("^uganda","Africa") %>%
           str_replace("^united arab emirates","Asia") %>%
           str_replace("^united kindom|united kingdom.*","Europe") %>%
           str_replace("^uruguay","South America") %>%
           str_replace("^vietnam","Asia") %>%
           str_replace("^zimbabwe","Africa") %>%
           str_replace("^uk.*|^unites kingdom|^wales|^wales Europe|^wales uk","Europe") %>%
           str_replace("^us.*|^usa.*|^virginia","North America") %>% 
           str_replace("^Europe uk*","Europe") %>%  
           str_replace("^Europe Europe*","Europe") %>%  
           str_replace("^South America but my org is in thailand*","Asia")) %>%
  filter(str_detect(Continent, "Europe|North America")) %>% 
  mutate(years_of_experience_in_field, 
         Experience = case_when(
           years_of_experience_in_field == "1 year or less" ~ "1<",
           years_of_experience_in_field == "11 - 20 years" ~ "11-20",
           years_of_experience_in_field == "2 - 4 years" ~ "2-4",
           years_of_experience_in_field == "21 - 30 years" ~ "21-30",
           years_of_experience_in_field == "31 - 40 years" ~ ">31",
           years_of_experience_in_field == "41 years or more" ~ ">31",
           years_of_experience_in_field == "5-7 years" ~ "5-7",
           years_of_experience_in_field == "8 - 10 years" ~ "8-10")) %>%
  mutate(Salary = case_when(
    Continent == "North America" & currency == "USD" ~ annual_salary/1,
    Continent == "North America" & currency == "CAD" ~ annual_salary/1.2,
    Continent == "Europe" & currency == "CHF" ~ annual_salary/0.89,
    Continent == "Europe" & currency == "GBP" ~ annual_salary/0.7,
    Continent == "Europe" & currency == "SEK" ~ annual_salary/8.29,
    Continent == "Europe" & currency == "EUR" ~ annual_salary/0.81)) %>%
  filter(Salary > 15000) %>%
  mutate(industry = str_replace(industry, "Government and Public Administration","Public Administration"))

tab <- survey_filtered %>% 
  select(Experience , gender, Continent, Salary) %>% 
  group_by(Experience,gender,Continent) %>% 
  filter(str_detect(gender, "Woman|Man")) %>%
  summarize(Median = median(Salary, na.rm=TRUE)) %>% 
  pivot_wider(names_from = c(gender,Continent),values_from = Median) 

  
kable(tab, booktabs = TRUE) %>%
  kable_styling(position = "center")

```


```{r}

cols <- c("Men"="#166FCC","Woman"="#FF4040")
fills <- c("EU"="#CD0074","US"="#FF7400")
p_median <- tab  %>% ggplot(aes(x = Experience, group = 1)) +
  geom_ribbon(aes(x = Experience, ymin = Woman_Europe, ymax = Man_Europe, fill = "EU"), alpha = 0.5) +
  geom_ribbon(aes(x = Experience, ymin = `Woman_North America`, ymax = `Man_North America`, fill = "US"), alpha = 0.5) +
  geom_line(aes(x = Experience, y = Man_Europe, group = "man", colour="Men")) +
  geom_line(aes(x = Experience, y = `Woman_North America`, group = "woman", colour="Woman")) +
  geom_line(aes(x = Experience, y = `Man_North America`, group = "man", colour="Men")) +
  geom_line(aes(x = Experience, y = Woman_Europe, group = "woman", colour="Woman")) +
  theme(legend.position="bottom",
    axis.title.y = element_text(size = 12, face = "plain"),
    axis.text = element_text(family = "Roboto Mono", size = 8.5),
    axis.ticks.x = element_line(colour = "grey40", size = 0.3),
    axis.ticks.y = element_line(colour = "transparent", size = 0.3),
    panel.grid.major.y = element_line(colour = "grey40", size = 0.3),
    panel.border = element_blank(),
    strip.background = element_rect(color = "transparent"),
    strip.text = element_text(size = 10),
    panel.spacing = unit(15, "pt"),
    legend.key	 = element_rect(color = "transparent")) +
  scale_x_discrete(limits = c("1<", "2-4", "5-7", "8-10",  "11-20", "21-30",  ">31")) +
  xlab("Experience in years") + 
  ylab("Annual salary") +
  scale_y_continuous(label = scales::unit_format(unit = "K", scale = 1e-3, sep = "")) + 
scale_colour_manual(name="Gender",values=cols, guide = guide_legend(override.aes=aes(fill=NA))) +
    scale_fill_manual(name="Gap",values=fills, guide_legend(override.aes=aes(fill=NA))) 

p_median
```

```{r}
tab_2 <- survey_filtered %>% 
  filter(Salary < 500000) %>%
  select(Experience , gender, Continent, Salary, industry) %>% 
  group_by(Experience, gender, Continent, industry) %>% 
  filter(str_detect(gender, "Woman|Man")) %>%
  summarize(Median = median(Salary, na.rm=TRUE))  %>%
  pivot_wider(names_from = c(gender, Continent),values_from = Median) %>% 
  drop_na() %>%
  filter(str_detect(industry, "Accounting, Banking & Finance|Business or Consulting|Computing or Tech|Education (Higher Education)|Public Administration|Health care|Media & Digital")) 
```


```{r}
p_facet <- tab_2 %>% ggplot(aes(x = Experience, y = Man_Europe, group = 1)) +
  geom_ribbon(aes(x = Experience, ymin = Woman_Europe, ymax = Man_Europe), fill = "#CD0074", alpha = 0.5) +
  geom_ribbon(aes(x = Experience, ymin = `Woman_North America`, ymax = `Man_North America`), fill = "#FF7400", alpha = 0.5) +
  geom_line(aes(x = Experience, y = Man_Europe, group = 1), color = "#166FCC") +
  geom_line(aes(x = Experience, y = `Woman_North America`, group = 1), color = "#FF4040") +
  geom_line(aes(x = Experience, y = `Man_North America`, group = 1), color = "#166FCC") +
  geom_line(aes(x = Experience, y = Woman_Europe, group = 1), color = "#FF4040") +
  theme(legend.position = "none",
        axis.title.y = element_text(size = 12, face = "plain"),
        axis.text = element_text(family = "Roboto Mono", size = 8.5),
        axis.ticks.x = element_line(colour = "grey40", size = 0.3),
        axis.ticks.y = element_line(colour = "transparent", size = 0.3),
        panel.grid.major.y = element_line(colour = "grey40", size = 0.3),
        panel.border = element_blank(),
        strip.background = element_rect(color = "transparent"),
        strip.text = element_text(size = 10),
        panel.spacing = unit(15, "pt")) +
  scale_x_discrete(limits = c("1<", "2-4", "5-7", "8-10",  "11-20", "21-30")) +
  xlab("Experience in years") + 
  ylab("Annual salary") +
  scale_y_continuous(label = scales::unit_format(unit = "K", scale = 1e-3, sep = "", expand = c(0, 0))) +
  facet_wrap(. ~ industry, nrow = 2) +
  labs(x = NULL, y = "\n", 
         caption = "\n\nVisualization by Vincent Manzanilla  |  Data source: Ask a Manager Survey")
```

```{r title}
## left-alligned title
p_title <- ggplot(data.frame(x = 1:2, y = 1:10)) +
  labs(x = NULL, y = NULL,
       title = "Gender gap in annual salary for managers",
       subtitle = "Self reported annual salaries for the US and Europe.\n") +
  theme(line = element_blank(),
        plot.background = element_rect(fill = "transparent", color = "transparent"),
        panel.background = element_rect(fill = "transparent"),
        panel.border = element_rect(color = "transparent"),
        axis.text = element_blank())
```

```{r full-panel, fig.width = 14, fig.height = 5.5}
 p_title + p_median +  p_facet + plot_layout(widths = c(0, 0.60, 1, 0)) +
    theme(legend.position='bottom')

ggsave("2021-01-18-Ask-a-Manager-Survey", width = 14, height = 5.6, device = pdf)
```


```{r}
sessionInfo()
```